On the Synchronization Power of the Openflow Data Plane

### 2015-03-29 inband-sync.tex
added sync file (a copy of main file)

%\usepackage{framed}
\usepackage{framed,color}
\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
\definecolor{mygreen}{rgb}{0,0.5,0}

@@

@@

\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
\subsection{The flow entry modification command (FlowMod)}
- Creating an entry with the same match part replaces the existing one.

\subsubsection{The CHECK\_OVERLAP flag}
- creating and entry with overlapping match (same priority) and different instructions fails.
\subsection{Bundling}
- failed command aborts the bundle.

@@

We assume transactional access to  two kinds of objects: a
\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
given (bounded or unbounded) set, and a read-write \emph{memory} abstraction
object, which we later use for storing the current used policy identifier.
%
\subsubsection{A resource claimer}
A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
\emph{unclaim} and \emph{\claimcheck} with the following sequential
semantics:
%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

@@

\subsubsection{A memory abstraction}
A memory can be accessed with the following sequential
semantics:

Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{abortIfUnequal}$ does).

@@

%features , including
%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
%[[PK let us leave for the moment
%We will describe later how to implement some of these features
%in older Openflow versions where they are not
%supported. \textbf{TODO}
%]]
Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, $\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
and $\textit{\cas}(x)$, respectively. 
Algorithm~\ref{alg:commit} describes the procedure executed by
a transaction at \emph{commit time}.

@@

In order to check if a number is claimed we try to create the corresponding entry while setting the
flag \textsf{CHECK\_OVERLAP} and using different action, thereby inflicting a failure in case an entry with overlapping match value exists. If the check succeeds we delete the entry. Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$. Therefore if one is used in \claimcheck after the other used in claim, a conflict is detected iff $x_1=x_2$, as expected.

Note that checking for claims must contain a deletion of the added flow entry otherwise a successive check operation might mistakenly fail. Moreover the two commands (adding and deleting of the flow entries) must be sent as part of the same bundle to support concurrent checks by different controllers.

@@

\begin{algorithm}[t]
    \caption{$\textit{claim}(x)$}
    \label{alg:claim}
    \begin{algorithmic}[1]
    \Require \emph{self} as the calling controller id.
%    \Ensure
    		\State $value \gets x\concat x$
    		\State $mask \gets self\concat all-1$
	    	\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets x\concat self$
    		\State $flag \gets 0$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{$\textit{\claimcheck}(x)$}
    \label{alg:check}
    \begin{algorithmic}[1]
    \Require \emph{self} as the calling controller id.
%    \Ensure
    		\State $value \gets x\concat x$
    		\State $mask \gets self\concat all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets x\concat self$
    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
			\Return $cmd1$,$cmd2$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{$\textit{unclaim}(x)$}
    \label{alg:unclaim}
    \begin{algorithmic}[1]
    \Require \emph{self} as the calling controller id.
%    \Ensure
    		\State $value \gets x\concat x$
    		\State $mask \gets self\concat all-1$
    		%\State $cookie \gets x\concat self$
    		\State $match \gets (value,mask)$
    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE, cookie) $

@@

We use similar techniques for implementing memory write and \compare 
 operations.
 As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
  the match part of the added flow entry is used as the memory address and the action as the written value. Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in casae they share the same match parts (as described in Section \ref{sec:background}).

In order to check if a value is written in a givven address, we try to add it the same way as in the write operation  but we also set the \textsf{CHECK\_OVERLAP} flag. If the value is not there, our action differ from the action of the existsing entry thereby inflicting a failure.
In case the value is there, our flow entry replaces the existing flow entry with identical one.
In addition we present in Algorithm \ref{alg:cas} a simple implementation of Comapare and set (CAS) usin the two last operations.

\begin{algorithm}[t]
    \caption{$\textit{write}(addr,k)$}
    \label{alg:write}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure
    		\State $value \gets addr$
    		\State $mask \gets  all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets k$
    		\State $flag \gets 0$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{$\textit{\compare}(addr,k)$}
    \label{alg:compare}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure
    		\State $value \gets addr$
    		\State $mask \gets  all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets k$
    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{$\textit{\cas}(addr, old,new)$}
    \label{alg:cas}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure

@@

\begin{algorithm}[t]
    \caption{$\textit{try-commit}()$}
    \label{alg:commit}
    \begin{algorithmic}[1]
    \Require  current bundle id $bid$.
		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
    		\State  $cmd.bundle\_id \gets bid$
    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST} 
    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
    		\State send $cmd$
		\Return $cmd$
    \end{algorithmic}
\end{algorithm}

>>>>

%\usepackage{framed}
\usepackage{framed,color}
\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
\definecolor{mygreen}{rgb}{0,0.5,0}
\definecolor{mygrey}{rgb}{0.5,0.5,0.5}

@@

\usepackage{listings}

@@

\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
\subsection{The flow entry modification command (FlowMod)}
\lstset{% general command to set parameter(s)
	basicstyle=\small, % print whole listing small
	keywordstyle=\color{black}\bfseries, % underlined bold black keywords
	identifierstyle=, % nothing happens
	commentstyle=\color{mygrey}, % white comments
	stringstyle=\ttfamily, % typewriter type for strings
	showstringspaces=false, % no special string spaces
	language=C}
\begin{lstlisting}
struct ofp_flow_mod {
	struct ofp_header header;
	uint64_t cookie; /* Opaque controller-issued identifier. */
	uint64_t cookie_mask; /* Mask used to restrict the cookie bits
		that must match when the command is
		OFPFC_MODIFY* or OFPFC_DELETE*. A value
		of 0 indicates no restriction. */
	/* Flow actions. */
	uint8_t table_id; /* ID of the table to put the flow in.
		For OFPFC_DELETE_* commands, OFPTT_ALL
		can also be used to delete matching
		flows from all tables. */
	uint8_t command; /* One of OFPFC_*. */
	uint16_t idle_timeout; /* Idle time before discarding (seconds). */
	uint16_t hard_timeout; /* Max time before discarding (seconds). */
	uint16_t priority; /* Priority level of flow entry. */
	uint32_t buffer_id; /* Buffered packet to apply to, or
		OFP_NO_BUFFER.
		Not meaningful for OFPFC_DELETE*. */
	uint32_t out_port; /* For OFPFC_DELETE* commands, require
		matching entries to include this as an
		output port. A value of OFPP_ANY
		indicates no restriction. */
	uint32_t out_group; /* For OFPFC_DELETE* commands, require
		matching entries to include this as an
		output group. A value of OFPG_ANY
		indicates no restriction. */
	uint16_t flags; /* Bitmap of OFPFF_* flags. */
	uint16_t importance; /* Eviction precedence (optional). */
	struct ofp_match match; /* Fields to match. Variable size. */
	/* The variable size and padded match is always followed by instructions. */
	//struct ofp_instruction_header instructions[0];
		/* Instruction set - 0 or more. The length
		of the instruction set is inferred from
		the length field in the header. */
};
\end{lstlisting}

- Creating an entry with the same match part replaces the existing one.

\subsubsection{The CHECK\_OVERLAP flag}
- creating and entry with overlapping match (same priority) and different instructions fails.
\subsection{Bundling}
- failed command aborts the bundle.
"When the commit fails, the switch must generate an error message corresponding to the
message that could not be applied. The error message must use the xid of the offending message,
the error data field corresponding to that message and the error code corresponding to the failure. If
multiple messages are generating errors, the switch may return only the first error found or generate
multiple error messages for the same bundle."
\subsubsection{Bundle control messages}

%/* Message structure for OFPT_BUNDLE_CONTROL. */
\begin{lstlisting}
struct ofp_bundle_ctrl_msg {
	struct ofp_header header;
	uint32_t bundle_id; /* Identify the bundle. */
	uint16_t type; /* OFPBCT_*. */
	uint16_t flags; /* Bitmap of OFPBF_* flags. */
	/* Bundle Property list. */
	struct ofp_bundle_prop_header properties[0]; /* Zero or more properties. */
};
\end{lstlisting}
\subsubsection{Bundle add message}

%/* Message structure for OFPT_BUNDLE_ADD_MESSAGE.
\begin{lstlisting}
* Adding a message in a bundle is done with. */
struct ofp_bundle_add_msg {
	struct ofp_header header;
	uint32_t bundle_id; /* Identify the bundle. */
	uint16_t pad; /* Align to 64 bits. */
	uint16_t flags; /* Bitmap of OFPBF_* flags. */
	struct ofp_header message; /* Message added to the bundle. */
	/* If there is one property or more, 'message' is followed by:
	* - Exactly (message.length + 7)/8*8 - (message.length) (between 0 and 7)
	* bytes of all-zero bytes */
	/* Bundle Property list. */
	//struct ofp_bundle_prop_header properties[0]; /* Zero or more properties. */
};
\end{lstlisting}

@@

We assume transactional access to  two kinds of objects: a read-write \emph{memory} abstraction
object, which we later use for storing the current used policy identifier, and a
\emph{resource-claimer} object, provides claiming a resource id (e.g. policy \emph{identifier}) in a
given (bounded or unbounded) set.
In theory one can use the memory abstraction to implement the resource claimer, however we found it far more simple and efficient to consider the resource claimer as base object.
%

A memory can be accessed with the following sequential
semantics:

Note that reading the memory can be done by just reading the switch configuration however such action doesn't affect the execution of the transaction (as $\textit{\compare}$ does).

A resource-claimer object implements a weak type of locking, exports operations \emph{claim},
\emph{unclaim} and \emph{\claimcheck} with the following sequential
semantics:
%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

@@

@@

\subsubsection{A resource claimer}
%features , including
%\emph{bundling}, \emph{conditional application}, and \emph{groups}.
%[[PK let us leave for the moment
%We will describe later how to implement some of these features
%in older Openflow versions where they are not
%supported. \textbf{TODO}
%]]
Algorithms~\ref{alg:claim}-\ref{alg:cas} describe implementations of  
transactional operations  $\textit{claim}(x)$, $\textit{\claimcheck}(x)$, $\textit{unclaim}(x)$, $\textit{write}(x)$,  $\textit{\compare}(x)$
and $\textit{\cas}(x)$, respectively. 
Algorithm~\ref{alg:commit} describes the procedure executed by
a transaction at \emph{commit time}.

@@

In order to check if a number is claimed we try to create the corresponding entry while setting the
flag \textsf{CHECK\_OVERLAP} and using different action ($2$ instead of $1$), thereby inflicting a failure in case an entry with overlapping match value exists. If the check succeeds we delete the entry. Considering $m_1$ and $m_2$ defined above and their respective actions $a_1=x_1\concat c1$ and $a_2=x_2\concat c2$,  they overlap with different actions iff $x_1=x_2$. Therefore if one is used in \claimcheck after the other used in claim, a conflict is detected iff $x_1=x_2$, as expected.

Note that for switch space reasons, checking for claims contains also a deletion of the added tester flow entry but it has no affect on claims and checks of other controllers and therefore the deletion command doesn't have to be sent as part of a bundle. This makes the \textsf{\claimcheck} operation atomic by nature, similar to the \textsf{claim} and \textsf{unclaim}, as being implemented by essentially one command.

@@

\begin{algorithm}[H]
    \caption{$\textit{claim}(x)$}
    \label{alg:claim}
    \begin{algorithmic}[1]
    \Require \emph{self} as the calling controller id.
%    \Ensure
    		\State $value \gets x\concat x$
    		\State $mask \gets self\concat all-1$
	    	\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets 1$
    		\State $flag \gets 0$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \caption{$\textit{\claimcheck}(x)$}
    \label{alg:check}
    \begin{algorithmic}[1]
    \Require \emph{self} as the calling controller id.
%    \Ensure
    		\State $value \gets x\concat x$
    		\State $mask \gets self\concat all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets 2$
    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
    		\State $cmd1\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
    		\State $cmd2\gets \textsc{FlowMod}(match, op = DELETE) $
			\Return $cmd1$,$cmd2$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \caption{$\textit{unclaim}(x)$}
    \label{alg:unclaim}
    \begin{algorithmic}[1]
    \Require \emph{self} as the calling controller id.
%    \Ensure
    		\State $value \gets x\concat x$
    		\State $mask \gets self\concat all-1$
    		%\State $cookie \gets x\concat self$
    		\State $match \gets (value,mask)$
    		\State $cmd\gets \textsc{FlowMod}(match, op=DELETE) $

@@

\subsubsection{A memory abstraction}

We use similar techniques for implementing memory \memwrite and \compare 
 operations.
 As can be seen in Algorithms \ref{alg:write}-\ref{alg:compare}, 
  the match part of the added flow entry is used as the memory address and the action as the written value. Values written to the same address will replace one another, following the command definition to replace old flow entry with a new one in case they share the same match parts (as described in Section \ref{sec:background}).

In order to check if a value is written in a given address, we try to add it the same way as in the \memwrite operation  but we also set the \textsf{CHECK\_OVERLAP} flag. If the value is not there, our action differ from the action of the existing entry thereby inflicting a failure.
In case the value is there, our flow entry replaces the existing flow entry with identical one.
In addition we present in Algorithm \ref{alg:cas} a simple implementation of Compare and set (CAS) using the two last operations.

\begin{algorithm}[h]
    \caption{$\textit{write}(addr,k)$}
    \label{alg:write}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure
    		\State $value \gets addr$
    		\State $mask \gets  all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets k$
    		\State $flag \gets 0$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
    \caption{$\textit{\compare}(addr,k)$}
    \label{alg:compare}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure
    		\State $value \gets addr$
    		\State $mask \gets  all-1$
    		\State $match \gets (value,mask)$
    		%\State $cookie \gets x\concat self$
    		\State $action \gets k$
    		\State $flag \gets \textsf{OFPFF\_CHECK\_OVERLAP}$
    		\State $cmd\gets \textsc{FlowMod}(match, op = ADD, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
    \caption{$\textit{\cas}(addr, old,new)$}
    \label{alg:cas}
    \begin{algorithmic}[1]
    %\Require  id $i$.
%    \Ensure

@@

\begin{algorithm}[h]
    \caption{$\textit{try-commit}()$}
    \label{alg:commit}
    \begin{algorithmic}[1]
    \Require  current bundle id $bid$.
		    \State $cmd \gets struct ofp_bundle_ctrl_msg$
    		\State  $cmd.bundle\_id \gets bid$
    		\State  $cmd.type \gets $ \textsf{OFPBCT\_COMMIT\_REQUEST} 
    		\State  $cmd.flags \gets $ \textsf{OFPBF\_ATOMIC}
    		\State send $cmd$
		\Return $cmd$
    \end{algorithmic}
\end{algorithm}
